Air cargo load and route planning in pickup and delivery operations

### 2023-06-07 CAIE/results.tex
Experiments continues

In each $scenario$\/, we indicate in bold the best value of $f$\/ found. In each table, we also indicate in bold the best {\bf Normalized}\/ and {\bf Speed-up}\/ values.
{\bf run time (s)} is the average time to solve a problem instance.

As the MIP solver was not capable of solving all scenarios in polynomial time or reached the computer's Read-Access Memory (RAM, 16 GiB) too early, we decided to set the Gurobi {\it MIPgap} parameter to 1\% to control the minimal quality of the returned solutions and abbreviate the solution time.
For more details on the Gurobi {\it MIPgap}, refer to \cite{mipgap}.

@@

97\% of 16GiB was the RAM state before we aborted the Gurobi execution on solving scenario 5 in Table \ref{tab:20}.

@@

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\color{violet}$surplus=2.0$} and 3600s as time limit.}  \label{tab:100}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\bf Tours} & $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ \specialcell{$\pi_{TSP_1}$\\ $\pi_{TSP_2}$ }} & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 24.39 & 26.46 & 27.09 & 24.00 & 1.00 \\ % 119.69
		&&                                                                               {\bf run time (s)}  & 162   & 95    & 74    & 68    & 67    & 1.0  \\ % 466
		\cmidrule(lr){2-9}	
		&                                                                   \multirow{2}{*}{\it Shims} & $f$ & 16.48 & 24.06 & 26.15 & 26.25 & 23.54 &  0.97 \\ % 116,48
		&&                                                                                {\bf run time (s)} & 5     & 6     & 8     & 10    & 11    &  11.7  \\ % 40

@@

		                                 \multirow{4}{*}{\it $S_{K}$ } & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 25.24 & 26.46 & 29.11 &     &  \\ % 
		&&                                                                                {\bf run time (s)} & 151   & 123   & 319   & 689   &     &   \\ % 
		\cmidrule(lr){2-9}	
		&                                                                \multirow{2}{*}{\it Shims}    & $f$ & 16.35 & 24.30 & 26.24 & 28.65 & 28.74   &   \\ % 
		&&                                                                                {\bf run time (s)} & 5     & 9     & 37    & 108   & 625     &   \\ % 
		\bottomrule

@@

@@

\begin{table}[H]
\centering
\caption{Solutions with volume {\color{red}$surplus=1.2$} with different time limitations for $S_K$ tours.}  \label{tab:20time}
\footnotesize
\begin{tabular}{ccccccccc}
\toprule
  $method$  & {\bf \specialcell{Time\\limit (s)}} & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}\\
\toprule

\multirow{6}{*}{\it Gurobi* }& \multirow{2}{*}{1200} & $f$ & 8.53 &    &    &    & x   &  \\ % 
                             &        &  {\bf run time (s)}& 29   &       &       &       & x   &   \\ % 
\cmidrule(lr){2-9}                              
                             & \multirow{2}{*}{2400}& $f$  & 8.53 & 12.25 & 13.31 & 14.59  & 13.59   &  \\ % 
                             &        &  {\bf run time (s)}& 29   & 40    & 129   & 313    & 1460   &   \\ % 
\cmidrule(lr){2-9}                              		
                             & \multirow{2}{*}{3600}& $f$  & 8.51 & 12.25 & 13.32 & 14.61 & x   & 1.00 \\ % 4869
                             &        &  {\bf run time (s)}& 25   & 36    & 121   & 297   & x   & 1.0 \\ % 479		
\midrule
\multirow{6}{*}{\it Shims}   & \multirow{2}{*}{1200} & $f$ &    &    &    &    &    &   \\ % 
                             &        & {\bf run time (s)} &    &    &    &    &    &   \\ % 
\cmidrule(lr){2-9}                              
                             & \multirow{2}{*}{2400}& $f$  &    &    &    &    &    &  \\ % 
                             &        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 	
\cmidrule(lr){2-9}                              
                             & \multirow{2}{*}{3600} & $f$ & 8.11 & 12.18 & 13.25 & 14.57 & 14.33 & 0.99 \\ % 4811 
                             &        &  {\bf run time (s)}& 4    & 8     & 33    & 95    & 572   & 3.4  \\ % 140 	                                        
\bottomrule

@@

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\color{blue}$surplus=1.5$} with different time limitations for $S_K$ tours.}  \label{tab:50time}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		$method$  & {\bf \specialcell{Time\\limit (s)}} & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}\\
		\toprule

		\multirow{6}{*}{\it Gurobi* }& \multirow{2}{*}{1200} & $f$ &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$  &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              		
		& \multirow{2}{*}{3600}& $f$  & 11.81 & 17.01 & 18.25 & 20.59 & 18.36   & 1.00 \\ % 8602
		&        &  {\bf run time (s)}& 53    & 61    & 193   & 470   & 2241    & 1.0  \\ % 3018		
		\midrule
		\multirow{6}{*}{\it Shims}   & \multirow{2}{*}{1200} & $f$ &    &    &    &    &    &   \\ % 
		&        & {\bf run time (s)} &    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$  &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 	
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{3600} & $f$ & 11.33 & 16.97 & 17.56 & 20.43 & 17.94   & 0.98  \\ % 8423 
		&        &  {\bf run time (s)}& 4     & 8     & 34    & 96    & 591     & 4.1 \\ % 733 	                                        
		\bottomrule

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\color{violet}$surplus=2.0$} with different time limitations for $S_K$ tours.}  \label{tab:100time}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		$method$  & {\bf \specialcell{Time\\limit (s)}} & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}\\
		\toprule

		\multirow{6}{*}{\it Gurobi* }& \multirow{2}{*}{1200} & $f$ &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$  &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              		
		& \multirow{2}{*}{3600}& $f$  & 17.75 & 25.24 & 26.46 & 29.11 &    &  \\ % 
		&        &  {\bf run time (s)}& 126   & 103   & 266   & 574   &     &   \\ % 		
		\midrule
		\multirow{6}{*}{\it Shims}   & \multirow{2}{*}{1200} & $f$ &    &    &    &    &    &   \\ % 
		&        & {\bf run time (s)} &    &    &    &    &    &   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$  &    &    &    &    &    &  \\ % 
		&        &  {\bf run time (s)}&    &    &    &    &    &   \\ % 	
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{3600} & $f$ & 8.11    & 12.18   & 13.25   & 14.57  & 14.33    &  0.99 \\ % 4811 
		&        &  {\bf run time (s)}& 4       & 8       & 33      & 95     & 572      &  3.4  \\ % 140 	                                        
		\bottomrule

@@

\subsubsection{Proposed strategy assessment}

@@

\begin{table}[H]
\centering
\caption{{\it Shims} best and least cost tour results}  \label{tab:20least}
\footnotesize
\begin{tabular}{ccccccccc}
\toprule
{$surplus$}              & {\bf Tour} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4}  & {\bf 5}   & {\bf Sum} & {\bf Improvement}\\

\midrule[.1pt]	
\multirow{2}{*}{$1.2$}   & Least cost    & 7.83    & 11.34 & 12.77 & 13.20 & 12.22  & 57.36 &       \\ 
                         & Best          & 8.11    & 12.18 & 13.25 & 14.57 & 14.33  & 62.44 & 8.9\%  \\ 
\midrule[.1pt]	
\multirow{2}{*}{$1.5$}   & Least cost    & 10.89   & 16.65 & 17.80 & 18.69 & 16.45  & 80.48 & \\ 	
                         & Best          & 11.33   & 16.97 & 17.56 & 20.43 & 17.94  & 84.23 & 4.7\%  \\ 
\midrule[.1pt]	
\multirow{2}{*}{$2.0$}   & Least cost    &         & 24.06 & 26.15 & 26.25 & 23.54  &       &  \\ 
                         & Best          &         & 24.30 & 26.24 & 28.65 & 28.74  &       &  \\ 		                         
\midrule[.1pt]
 \multicolumn{8}{r}{{\bf Average improvement}}                                              & \% \\
\bottomrule
\end{tabular}
\normalsize
\end{table}

@@

Considering that the computer used in the experiments has a standard RAM consumption when idle of 2.5 GiB and all Gurobi executions consumed from 12 to 15 GiB of RAM, the Gurobi executions consumed 10 to 12.5 GiB in total.

All {\it Shims} executions consumed at most 1.5 GiB, or 3 to 4 GiB in total.

These details are important for a future application to adequately dimension the computer to be used.

In the beginning of this work, we devised, as one of the objectives, a method that may be run on a handheld computer. If this is the case, the use of a MIP solver may not be appropriate, as most handheld computers come with 4 GB of RAM.

>>>>

In each $scenario$\/, we indicate in bold the best value of $f$\/ found. In each table, we also indicate in bold the best {\bf Normalized}\/ and {\bf Speed-up}\/ values.

{\color{blue}
{\bf run time (s)} is the average time to solve a problem instance.

As the MIP solver was not capable of solving all scenarios in polynomial time or reached the computer's Read-Access Memory (RAM, 16 GiB) too early, we decided to set the Gurobi {\it MIPgap} parameter to 1\% to control the minimal quality of the returned solutions and abbreviate the solution time.
From now on, we will refer to this Gurobi-specific model as {\it Gurobi* }.
For more details on the Gurobi {\it MIPgap}, refer to \cite{mipgap}.

@@

97\% of 16GiB was the RAM state before we aborted the {\it Gurobi*} execution on solving scenario 5 in Table \ref{tab:20}.

@@

\begin{table}[H]
	\centering
	\caption{Solutions with volume {\color{violet}$surplus=2.0$} and 3600s as time limit.}  \label{tab:100}
	\footnotesize
	\begin{tabular}{ccccccccc}
		\toprule
		{\bf Tours} & $method$          & {\bf Scenarios} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4} & {\bf 5} & \specialcell{{\bf Normalized}\\{\bf Speed-up}}   \\
		\toprule
		\multirow{4}{*}{ \specialcell{$\pi_{TSP_1}$\\ $\pi_{TSP_2}$ }} & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 24.39 & 26.46 & 27.09 & 24.00 & 1.00 \\ % 119.69
		&&                                                                               {\bf run time (s)}  & 162   & 95    & 74    & 68    & 67    & 1.0  \\ % 466
		\cmidrule(lr){2-9}	
		&                                                                   \multirow{2}{*}{\it Shims} & $f$ & 16.73 & 24.06 & 26.15 & 26.25 & 23.54 &  0.97 \\ % 116,48
		&&                                                                                {\bf run time (s)} & 5     & 6     & 8     & 10    & 11    &  11.7  \\ % 40

@@

		                                 \multirow{4}{*}{\it $S_{K}$ } & \multirow{2}{*}{\it Gurobi* } & $f$ & 17.75 & 25.24 & 26.46 & 29.11 &  x   & 1.00 \\ % 9856
		&&                                                                                {\bf run time (s)} & 151   & 123   & 319   & 689   &  x   & 1.0  \\ % 1282
		\cmidrule(lr){2-9}	
		&                                                                \multirow{2}{*}{\it Shims}    & $f$ & 16.88 & 24.30 & 26.24 & 28.65 & 28.74 & 0.97  \\ % 9607
		&&                                                                                {\bf run time (s)} & 5     & 9     & 37    & 108   & 625   & 8.1  \\ % 159
		\bottomrule

@@

{\color{blue}
{\it Gurobi*} could not solve scenario 5 in Table \ref{tab:100} because $3600/K! = 0.71s$, not enough time for {\it Gurobi*} to find a feasible node solution.
}

@@

\begin{table}[H]
	\centering
	\caption{Solutions with different time limitations for $S_K$ tours.}  \label{tab:100time}
	\footnotesize
	\begin{tabular}{ccc|cc|cc|cc}
		\toprule
		&             \multicolumn{2}{r}{\bf Volume $surplus$} &\multicolumn{2}{c}{$1.2$}&\multicolumn{2}{c}{$1.5$}&\multicolumn{2}{c}{$2.0$} \\
		\midrule
$method$& {\bf \specialcell{Time\\limit (s)}} & {\bf Scenarios}    & {\bf 4} & {\bf 5} & {\bf 4} & {\bf 5} & {\bf 4} & {\bf 5} \\
		\toprule

		\multirow{6}{*}{\it Gurobi* }& \multirow{2}{*}{1200} & $f$ & 14.67   & x       &  20.64  &  x      & 27.97   & x \\ % 
		&        &  {\bf run time (s)}                             & 217     & x       &  304    &  x      & 579     & x \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$                               & 14.59   & 13.59   &  20.60  &  x      & 29.10   & x \\ % 
		&        &  {\bf run time (s)}                             & 313     & 1460    &  327    &  x      & 543     & x \\ % 
		\cmidrule(lr){2-9}                              		
		& \multirow{2}{*}{3600}& $f$                               & 14.61   & x       &  20.59  & 18.36   & 29.11   & x \\ % 
		&        &  {\bf run time (s)}                             & 297     & x       &  470    & 2241    & 689     & x \\ % 		
		\midrule
		\multirow{6}{*}{\it Shims}   & \multirow{2}{*}{1200} & $f$ & 15.14   &         &         &         &         &   \\ % 
		&        & {\bf run time (s)}                              & 73      &         &         &         &         &   \\ % 
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{2400}& $f$                               &         &         &         &         &         &   \\ % 
		&        &  {\bf run time (s)}                             &         &         &         &         &         &   \\ % 	
		\cmidrule(lr){2-9}                              
		& \multirow{2}{*}{3600} & $f$                              & 14.57   & 14.33   &  20.43  & 17.94   & 28.65   & 28.74 \\ %  
		&        &  {\bf run time (s)}                             &  95     &  572    &  96     & 591     & 108     &  625  \\ %  	                                        
		\bottomrule

@@

\begin{figure}[H]
\centering
\begin{tikzpicture}
	\begin{axis}[
		area style,
		]
		\addplot+[ybar interval,mark=no] plot coordinates { 
			(0,.3)
			(1,.5)
			(2,.6)
			(3,.5)
			(4,.25)
			(5,.1)
			(6,.5)
			(7,.25)
			(8,.1)
		};
	\end{axis}
\end{tikzpicture}
\captionof{figure}{ESTE GRÁFICO É APENAS UM MODELO}
\end{figure}

@@

\subsubsection{Proposed K! strategy assessment}

@@

\begin{table}[H]
\centering
\caption{{\it Shims} best and least cost tour results}  \label{tab:20least}
\footnotesize
\begin{tabular}{ccccccccc}
\toprule
{$surplus$}              & {\bf Tour} & {\bf 1} & {\bf 2} & {\bf 3} & {\bf 4}  & {\bf 5}   & {\bf Sum} & {\bf Improvement}\\
\toprule

\multirow{2}{*}{$1.2$}   & Least cost    & 7.83    & 11.34 & 12.77 & 13.20 & 12.22  & 57.36 &       \\ 
                         & Best in $K!$  & 8.11    & 12.18 & 13.25 & 14.57 & 14.33  & 62.44 & 8.9\%  \\ 
\midrule[.1pt]	
\multirow{2}{*}{$1.5$}   & Least cost    & 10.89   & 16.65 & {\color{red}17.80} & 18.69 & 16.45  & 80.48 & \\ 	
                         & Best in $K!$  & 11.33   & 16.97 & 17.56 & 20.43 & 17.94  & 84.23 & 4.8\%  \\ 
\midrule[.1pt]	
\multirow{2}{*}{$2.0$}   & Least cost    & 16.73   & 24.06 & 26.15 & 26.25 & 23.54  & 116.73&  \\ 
                         & Best in $K!$  & 16.88   & 24.30 & 26.24 & 28.65 & 28.74  & 124.81& 6.7\%   \\ 		                         
\midrule[.1pt]
\multicolumn{8}{r}{{\bf K! strategy advantage}}                                              & {\bf 6.8}\% \\

\bottomrule
\end{tabular}
\normalsize
\end{table}

@@

Most of the {\it Gurobi*} executions consumed over 12 GiB in total, and considering that the computer used in the experiments has a standard RAM consumption when idle of 3.5 GiB (with the Python Interface Development Environment (IDE), a PDF viewer, a system monitor, and a {\it LaTeX} IDE simultaneously open), the actual RAM consumption of {\it Gurobi*} was over 8.5 GiB.

All {\it Shims} executions consumed at most 1.5 GiB, requiring 5 GiB of total memory.
These details are important for a future application to adequately dimension the computer to be used.
